---
title: "exam2016"
output: github_document
---
## null hypothesis
*mean oral temperature does not have an effect on heart rate on male and females
## alternative hypothesis
*mean oral temperature has an effect on heart rate on male and females
## type of data being analysed
*
```{r}

```


```{r question1, echo=TRUE}
G <- read.csv("file:///C:/Users/Wits-User/Desktop/Exam2016/question1.csv")
library(tidyr)
# Tidy the data
head (G)
G <- xtabs(~body_temperature+male + female, data= G)
barplot(G, ylab = 'body_temperature', xlab = 'male and female', main = 'body temperature in male and females', col= c("pink","blue","red"))

```

## GitHub Documents

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

## Including Code

You can include R code in the document as follows:

```{r question 2, echo=TRUE}
m <- read.csv("file:///C:/Users/Wits-User/Desktop/Exam2016/question2.csv")
# tidy data
m <- xtabs(~handedness + first_stumble, data = m)
head(m)
# explore the data
dim(m)
names(m)
head(m)
tail(m)
summary(m)

#plot data
barplot(m, beside = TRUE, ylab= "intoxication", xlab = "balance", main = "the effect of drinking on handeness and ataxic walking", col= c("pink", "blue")



```

## Including Plots

You can also embed plots, for example:
## null hypothesis

* running time does not affect the number of calories consumed
## alternative hypothesis

* running time has an effect on the number of calories consumed
## type of data being analysed
* numerical data : continous data
## statistical test
* linear regression
*the generalized linear model because the diagnostics I ran were not normally distributed.
## assumption
* there is a trend bewteen x and y
* the residuals are normally distributed and Homoskedastcity
```{r question 3, echo=TRUE}
x <- read.csv("file:///C:/Users/Wits-User/Desktop/Exam2016/question3.csv")
# TIDY DATA
time= x$time
calories= x$calories
head(cbind(time, calories))
# scatter plot
plot(time, calories, xlab = "time", ylab = "calories")
abline( lm(d$calories~d$time, data = d), col="blue", lwd=2)
# linear regression
 x1 <-lm(x$time~x$calories, data= x)
 summary(x1)
 #diagnostic plot 1: Homoskedasticity
 plot( x = x1$fitted.values, y = x1$residuals, main = "Homoskedasticity", col="green")
 abline(h=0)
 #diagnostic plot 2: gaussian residual distribution
 qqnorm(x1$residuals)
 qqline(x1$residuals)
 # variables have non-gaussian error of distribution
 glm(x$calories~x$time, data= x)
 summary(x1)

```
## statistical results
*p = 7.95e-12 which is less than 0.05 therefore reject the null hypothesis.
* running has an effect on the consumption of calories
```{r}




#plot the data

barplot(x1, beside = TRUE, ylab = 'intoxication', xlab= 'presentation', main = 'effect of intoxication on handedness and ataxia', col= c("pink", "turquoise"))

legend('top', c("handedness", "ataxia"), fill = c("pink", "turquoise"))

```

